---
title: "CyboRg 2021-10 (October)"
output: 
  word_document: 
    fig_width: 5
    fig_height: 5
---

# CyboRg session 2021-10-05

# 1. Regular grid inside irregular polygon, with user-defined spacing and offset

This is code to populate an irregular polygon with a rectangular grid of points with user-selected spacing. It represents part of the first phase of development of an R package to generate different sampling designs given a site boundary and number or spacing of samples. The next code chunks (# 2.) also populate the irregular shape with a grid, but this time with user-selection of the number of samples.

## set user-defined parameters
```{r}
boundfile <- "sv_boundary.csv"
spacing <- 40 # spacing in metres for UTM data
offset <- 25 # x,y offset in metres for UTM data
```

## read data and set up basic boundary parameters
```{r}
library(sp)
irreg <- read.csv(boundfile)
str(irreg)
# make a spatial polygon (package 'sp') to calculate area
irregpoly <- Polygon(cbind(irreg$Easting, irreg$Northing), hole=F)
str(irregpoly)
# use floor, ceiling to get integer values for boundary extremes
extremes <- as.integer(c(min(floor(irreg$Easting)),
              min(floor(irreg$Northing)),
              max(ceiling(irreg$Easting)),
              max(ceiling(irreg$Northing))))
# two useful areas
area <- irregpoly@area
maxarea <- (extremes[3]-extremes[1]) * (extremes[4]-extremes[2])
```

## use boundary data to set up and generate square grid
```{r message=FALSE, warning=FALSE}
# we need the geoR package for the locations.inside() function
library(geoR)
grid_max <- expand.grid(seq(extremes[1], 
                            extremes[3], by = spacing) + offset,
                        seq(extremes[2], 
                            extremes[4], by = spacing) + offset)
inside <- locations.inside(locations = grid_max,
                          borders = irreg[, c("Easting", "Northing")])
str(inside)
cat("\nOriginal grid has", NROW(grid_max),"points, with",NROW(inside),"of these inside sampling area.\n")
```

## check how it looks with a plot

We use several objects from the previous 2 chunks.

```{r fig.height=7, fig.width=5}
par(mfrow = c(1,1), mar = c(3,3,1,1), mgp = c(1.6, 0.3, 0), 
    tcl = 0.25, font.lab = 2, lend=1, ljoin=2, cex.axis = 0.8)
with(irreg, plot(Easting, Northing, asp = 1, type = "l"))
rect(extremes[1], extremes[2], 
     extremes[3], extremes[4], 
     border = "chocolate", lty = 2)
points(inside, col = "purple", pch = 3, cex = 0.8)

legend("bottomleft", inset = 0.05, y.intersp = 1.2, bty = "o", box.col = 8,
       legend = c(paste0("Spacing = ",spacing,"m, offset = ", offset,"m"),
                  paste("Irregular polygon",signif(area,5),"m\u00B2"), 
                  paste("Max. extent",signif(maxarea,6),"m\u00B2")),
       pch = c(3,NA,NA), col = c("purple",1, "chocolate"), lty = c(NA,1,2))
mtext(paste("Polygon has",length(inside[,1]),"points from original grid of",length(grid_max[,1]),"points"),
      3,-1)
```

## set user-defined parameters
```{r}
boundfile <- "sv_boundary.csv"
n <- 75 # desired number of points in grid
offset <- 5 # x,y offset in metres for UTM data
```

## read data and set up basic boundary parameters
```{r}
library(sp)
irreg <- read.csv(boundfile)
str(irreg)
# make a spatial polygon (package 'sp') to calculate area
irregpoly <- Polygon(cbind(irreg$Easting, irreg$Northing), hole=F)
str(irregpoly)
# use floor, ceiling to get integer values for boundary extremes
extremes <- as.integer(c(min(floor(irreg$Easting)),
              min(floor(irreg$Northing)),
              max(ceiling(irreg$Easting)),
              max(ceiling(irreg$Northing))))
# two useful areas
area <- irregpoly@area
maxarea <- (extremes[3]-extremes[1]) * (extremes[4]-extremes[2])
```

## use number of points, area ratio, and aspect ratio to set up and generate square grid
```{r message=FALSE, warning=FALSE}
# we need the geoR package for the locations.inside() function
library(geoR)
yxratio <- diff(range(extremes[c(2,4)]))/diff(range(extremes[c(1,3)]))
arearatio <- area/maxarea
# next 2 lines from some algebra done outside the code
# essentially we need to solve x*y*arearatio = n, where y/x = yxratio
# first to calculate number of points per side...
xside <- sqrt(n/(arearatio*yxratio))
yside <- yxratio*sqrt(n/(arearatio*yxratio))
# ...then use this and maxarea dimensions to calculate (rounded) spacing
spacing <- round(mean(diff(range(extremes[c(1,3)]))/xside,
              diff(range(extremes[c(2,4)]))/yside), 0)
grid_max <- expand.grid(seq(extremes[1], 
                            extremes[3], by = spacing) + offset,
                        seq(extremes[2], 
                            extremes[4], by = spacing) + offset)
# then find just the locations inside the grid
# which should be approx. arearatio * (points in grid_max)
inside <- locations.inside(locations = grid_max,
                          borders = irreg[, c("Easting", "Northing")])
str(inside)
cat("\nOriginal grid has", NROW(grid_max),"points, with",NROW(inside),"of these inside sampling area.\n",
    "(Note that actual and desired number of points may differ due to rounding.)\n")
```

## check how it looks with a plot

We use several objects from the previous 2 chunks.

```{r fig.height=7, fig.width=5.5}
par(mfrow = c(1,1), mar = c(3,3,1,1), mgp = c(1.6, 0.3, 0), 
    tcl = 0.25, font.lab = 2, lend=1, ljoin=2, cex.axis = 0.8)
with(irreg, plot(Easting, Northing, asp = 1, type = "l"))
rect(extremes[1], extremes[2], 
     extremes[3], extremes[4], 
     border = "chocolate", lty = 2)
points(inside, col = "darkorchid", pch = 10, cex = 1)

legend("bottomleft", inset = 0.05, y.intersp = 1.2, bty = "o", box.col = 8,
       legend = c(paste0("Calculated spacing = ",spacing,"m"), paste0("User-defined offset = ", offset,"m"),
                  paste("Irregular polygon",signif(area,5),"m\u00B2"), 
                  paste("Max. extent",signif(maxarea,6),"m\u00B2")),
       pch = c(10,NA,NA,NA), col = c("darkorchid",NA,1, "chocolate"), lty = c(NA,NA,1,2))
mtext(paste0("Polygon has ",NROW(inside)," points (target ",n,") from grid of ",NROW(grid_max)," points"),
      3,-1)
```

# 2. Making the code into a function

Here's some information about the rectBySpacing function in R Help format:

*Description*
A function to populate irregular polygon with rectangular grid points at defined x (and optionally y) spacings

*Usage*
rectBySpacing(boundary, xspacing = 0, yspacing = xspacing, xoffset = 0,
              yoffset = xoffset, proj4 = NULL)

*Arguments*
boundary    a matrix or data frame with the coordinates of the irregular polygon 
            boundary in the first two columns (x in 1st, y in 2nd)

xspacing    the spacing of the grid in the x-dimension (e.g. easting, longitude),
            with the same units as the coordinates (defaults to zero)

yspacing    spacing of the grid in the y-dimension (e.g. northing, latitude),
            with the same units as the coordinates (default is xspacing value)        

xoffset     the starting x-coordinate of the grid relative to the minimum x in 
            the boundary coordinates and with the same units (defaults to zero)

yoffset     the starting y-coordinate of the grid relative to the minimum y in 
            the boundary coordinates and with the same units (defaults to zero)

proj4       the proj4 string for the coordinate reference system (defaults to 
            "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs", i.e. EPSG:4326)

*Value*
An object of class "list" containing:

points_in   a 2-column data frame containing the coordinates of the points 
            inside the irregular polygon

x_spacing   a single numeric value of the x-spacing specified in the function call

y_spacing   a single numeric value of the y-spacing specified in the function call

x_offset    a single numeric value of the x-offset specified in the function call

y_offset    a single numeric value of the y-offset specified in the function call

npoints     the number of points inside the irregular polygon (integer)

max_points  the number of points in the grid constrained by a rectangle defined by 
            the minima and maxima of the irregular polygon coordinates (integer)

shape_area  a single numeric value of the area (in coordinate units squared) of 
            the irregular polygon

max_area    a single numeric value of the area (in coordinate units squared) of 
            the rectangle defined by the minima and maxima of the irregular polygon

extremes    a numeric vector containing min(x), min(y), max(x), max(y)

proj_4_str  a character string of class "CRS" [package "sp"] which is the 
            proj4 string for the coordinate reference system

## note that the function code uses the over() function from the sp package, rather than the locations.inside() function from the geoR package (which is based on the sp::over function anyway)

```{r}
rectBySpacing <-
  function(boundary,
           xspacing = 0,
           yspacing = xspacing,
           xoffset = 0,
           yoffset = xoffset,
           proj4 = NULL
  ) {
    require(sp)
    
    # if projection not defined, set to EPSG:4326 (LongLat)
    if (is.null(proj4)) {
      p4 <-
        CRS("+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")
    } else {
      p4 <- CRS(proj4)
    }
    
    # make boundary coords into SpatialPolygons object
    irregpoly <- Polygon(boundary[, c(1,2)], hole = F)
    irregPolys = Polygons(list(irregpoly),1)
    irregSpatialPolys = SpatialPolygons(list(irregPolys), 
                          proj4string = p4)
    
    # define limirs of initial (maximal) grid
    extremes <- c(min(floor(boundary[, 1])),
                  min(floor(boundary[, 2])),
                  max(ceiling(boundary[, 1])),
                  max(ceiling(boundary[, 2])))
    # calculate areas
    area <- irregpoly@area
    maxarea <-
      (extremes[3] - extremes[1]) * (extremes[4] - extremes[2])
    
    # make initial (maximal) grid
    grid_big <- expand.grid(
      seq(extremes[1],
          extremes[3], by = xspacing) + xoffset,
      seq(extremes[2],
          extremes[4], by = yspacing) + yoffset
    )
    # convert initial grid to SpatialPoints object
    gridSP <- SpatialPoints(grid_big, proj4string = p4)

    # user over() function from sp package to mask initial grid
    inOrOut <- as.vector(over(gridSP, irregSpatialPolys))
    insideSP <- gridSP[which(inOrOut>0)]
    # convert masked coordinates back into data frame
    inside <- as.data.frame(insideSP@coords)
    colnames(inside) <- c("x", "y")
    
    # make output object
  output <-
    list(
      points_in = inside,
      x_spacing = xspacing,
      y_spacing = yspacing,
      x_offset = xoffset,
      y_offset = yoffset,
      npoints = NROW(inside),
      max_points = NROW(grid_big),
      shape_area = area,
      max_area = maxarea,
      extremes = c(min(floor(boundary[, 1])),
                    min(floor(boundary[, 2])),
                    max(ceiling(boundary[, 1])),
                    max(ceiling(boundary[, 2]))),
      proj_4_str = p4
    )
  return(output)
  }
```

## testing the rectBySpacing() function - create an output object and plot

```{r fig.height=7, fig.width=6}
SVbound <- read.csv("sv_boundary.csv")
SVdesign <-
  rectBySpacing(SVbound,
                xspacing = 30,
                yspacing = 40,
                xoffset = 10,
                yoffset = 15,
                proj4 = "+proj=utm +zone=50 +south")
# -+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-
# check with a plot
par(
  mfrow = c(1, 1),
  mar = c(3, 3, 1, 1),
  mgp = c(1.6, 0.3, 0),
  tcl = 0.25,
  font.lab = 2,
  lend = 2,
  ljoin = 1,
  cex.axis = 0.8
)
with(SVbound, plot(Easting, Northing, asp = 1, type = "l"))
rect(SVdesign$extremes[1], SVdesign$extremes[2],
     SVdesign$extremes[3], SVdesign$extremes[4],
     border = "chocolate",
     lty = 2)
points(SVdesign$points_in,
       col = "purple",
       pch = 3,
       cex = 0.8)

legend(
  "bottomleft",
  inset = 0.04,
  y.intersp = 1.2,
  bty = "o",
  box.col = 8,
  title = expression(bold("Rectangular Grid Sampling Design")),
  legend = c(
    paste0("Spacing = ", SVdesign$x_spacing,"\u00D7",SVdesign$y_spacing), 
    paste0("  offset = ", SVdesign$x_offset,"\u00D7",SVdesign$y_offset),
    paste("Irregular polygon", signif(SVdesign$shape_area, 5), "m\u00B2"),
    paste("Max. extent", signif(SVdesign$max_area, 6), "m\u00B2")
  ),
  cex = 0.9, pch = c(3, NA, NA, NA),
  col = c("purple", NA, 1, "chocolate"),
  lty = c(NA, NA, 1, 2)
)
mtext(paste(
  "Polygon contains",
  SVdesign$npoints,
  "points from maximal grid of",
  SVdesign$max_points, "points"), 
  3, -1, col = "grey50", font = 3)
mtext("Charles Veryard and Smith's Lake Reserves, North Perth", 
  side = 3, line = 0.1, 
  col = "#27348B", font = 2, 
  cex = 0.9, family = "serif")
```

### tidy up
```{r}
rm(list = c("boundfile","spacing","offset","n","irreg","irregpoly",
            "extremes","area","maxarea","arearatio","yxratio",
            "xside","yside","grid_max","inside"))
```
