---
title: "CyboRg 2021-08 (August)"
output: 
  word_document: 
    fig_width: 5
    fig_height: 5
---
```{r read some data used in the next code chunk, paged.print=FALSE}
sv18 <- 
  read.csv("C:/Users/00028958/LocalData/R Projects/ENVT4461/sv2018_original.csv",
           stringsAsFactors=TRUE)
limits <- read.csv("limits.csv", stringsAsFactors = TRUE)
row.names(limits) <- c("As","Ba","Be","Cd","Cr","Cu","Mn","Ni","Pb","V","Zn")
print(limits)
```

# CyboRg 10 August 2021

## Function of the week - *which()*

An updated version of the code I showed in Cyborg20210810. I think I might have used the previous version to also show the t() (transpose) function.

```{r function of the week which(), paged.print=FALSE}
out <- 
  as.data.frame(cbind(as.numeric(as.character(sv18$Core[which(sv18$Cu > limits["Cu","EIL"])])),
      sv18$Cu[which(sv18$Cu > limits["Cu","EIL"])],
      sv18$Depth_upper[which(sv18$Cu > limits["Cu","EIL"])],
      sv18$Depth_lower[which(sv18$Cu > limits["Cu","EIL"])]))
colnames(out) <- c("Core","Cu (mg/kg)","DepthHi","DepthLo")
cat("Cores and depths in which Cu concentration > EIL\n");print(out)
out <- 
  as.data.frame(cbind(as.numeric(as.character(sv18$Core[which(sv18$Pb > limits["Pb","EIL"])])),
      sv18$Pb[which(sv18$Pb > limits["Pb","EIL"])],
      sv18$Depth_upper[which(sv18$Pb > limits["Pb","EIL"])],
      sv18$Depth_lower[which(sv18$Pb > limits["Pb","EIL"])]))
colnames(out) <- c("Core","Pb (mg/kg)","DepthHi","DepthLo")
cat("\nCores and depths in which Pb concentration > EIL\n");print(out) # \n represents a line break
out <- 
  as.data.frame(cbind(as.numeric(as.character(sv18$Core[which(sv18$Zn > limits["Zn","EIL"])])),
      sv18$Zn[which(sv18$Zn > limits["Zn","EIL"])],
      sv18$Depth_upper[which(sv18$Zn > limits["Zn","EIL"])],
      sv18$Depth_lower[which(sv18$Zn > limits["Zn","EIL"])]))
colnames(out) <- c("Core","Zn (mg/kg)","DepthHi","DepthLo")
cat("\nCores and depths in which Zn concentration > EIL\n");print(out) # \n represents a line break
rm(out) # remove temporary object(s)
```

# CyboRg 10 August 2021

## Function of the week - *(grep)*
```{r use grep to combine different names for the same thing}
# read the data #####
# downloaded from
# https://github.com/rfordatascience/tidytuesday/blob/master/data/2021/2021-08-17/computer.csv
computer <- read.csv("C:/Users/00028958/LocalData/R Projects/CyboRg/computer.csv")

# create a new empty column *new_char* full of NA values
computer$new_char <- rep(NA,nrow(computer))

# Combining different names for (probably) the same character ####
# find the row numbers where the character (column computer$char) 
# contains "Computer", put results into object x
x <- grep("Computer",computer$char)
# replace NA with "Computer" in the column new_char rows
computer$new_char[x] <- rep("Computer", length(x))
computer[x,c("char","new_char")] # show the results
```

# and something we talked about but didn't show

```{r Number of times a character says computer}
# lots of nested functions here!
# maybe we should also learn how to use pipes?
y <- as.data.frame(summary(as.factor(computer$char[grep("computer",computer$line)])))
colnames(y) <- "n"
y$Character <- row.names(y)
par(mar = c(7,4,1,1),mgp=c(1.5,0.2,0), las=2, cex.axis = 1, cex.lab = 1, 
        font.lab = 2, tcl=0.2)
barplot(y$n, names.arg = y$Character, cex.names = 0.8, 
        ylab="Number of times a\ncharacter says 'computer'",
        ylim = c(0,max(y$n)*1.05), col = "steelblue")
box()
```

# Week 3 of CyboRg - previews

## Flexible code
```{r How to identify just the numeric columns, paged.print=FALSE}
cat(rep("-",10),"All the column names",rep("-",10),"\n");names(computer)
num_cols <- as.logical(sapply(computer, is.numeric)) # Identify numeric columns
computer_num <- computer[,num_cols]
cat(rep("-",10),"Just the names of numeric columns",rep("-",10),"\n");names(computer_num)
cat(rep("-",20),"\n");head(computer_num)
cat(rep("-",20),"\n");cat("The dataset has",ncol(computer_num),"numeric columns\n")
```

```{r example of minimum-input loop for shapiro-wilk test, paged.print=FALSE}
# script authored by andrew rate (C) the university of western australia 2016-2021
# ratey.at.uwa@gmail.com
#
# this version creates new log- and power-transformed variables
# and tests all [un]transformed variables for normality
#
# load required packages
require(car)
# create temp object with names of variables to be transformed

# define starting and ending columns
c1 <- 10
cn <- 40 # probably not needed in this version 
subdata <- sv18[,c1:ncol(sv18)] # strip off first few columns (i.e. sample info)
n_cols <- as.logical(sapply(subdata, is.numeric)) # Identify numeric columns
numdata <- subdata[,n_cols]
rm(subdata)
names.of.cols <- names(numdata)
nvar <- length(names.of.cols)
#
# generate matrix of comma separated values
# and calculate new variables
#
# make initial output data frame
table <- data.frame("Variable"=rep(NA,nvar),
                             "W_orig"=rep(NA,nvar),
                             "p_orig"=rep(NA,nvar), "W_log_tr"=rep(NA,nvar),
                             "p_log_tr"=rep(NA,nvar), "W_pow_tr"=rep(NA,nvar),
                             "p_pow_tr"=rep(NA,nvar), "Pow_term"=rep(NA,nvar))
# start loop that assesses variable distributions and creates new variables
for (i in 1:nvar) {
  pt1 <- powerTransform(numdata[, i])
  sw0 <- shapiro.test(numdata[, i])
  sw1 <- shapiro.test(log10(numdata[, i]))
  sw2 <- shapiro.test((numdata[, i]) ^ as.vector(pt1$lambda))
  table[i,] <- c(names.of.cols[i], signif(sw0$statistic, 4), 
                                 signif(sw0$p.value, 4), signif(sw1$statistic, 4), 
                                 signif(sw1$p.value, 4), signif(sw2$statistic, 4),
                                 signif(sw2$p.value, 4), signif(as.vector(pt1$lambda), 4))
}
#
# output caption and table
# "\n" inserts a line break
{cat("\nTable. Shapiro-Wilk statistics and p-values for untransformed (_orig) and transformed
    (_log, _pow) variables from soil and sediment analysis at Ashfield Flats Reserve.\n\n")
print(table, row.names = FALSE)}
#
# export results to a csv file for Excel (if desired)
write.csv(table, file = "transformations.csv", row.names = FALSE)
write.table(table, "clipboard", sep="\t", row.names = F)

# remove temporary objects
# to keep R workspace tidy
rm(list=c("c1","cn","n_cols","numdata","nvar",
          "names.of.cols","pt1","sw0","sw1","sw2","i","table"))
# end code
```

```{r some code not used in the previous chunk}
  #numdata[paste0(names.of.cols[i],".log")]<-log10(numdata[i])
  # if ... else applies factor of -1 to 
  # power transforms with negative terms
  # if (as.vector(pt1$lambda) > 0) {
    #numdata[paste0(names.of.cols[i], ".pow")] <-
    # numdata[i] ^ as.numeric(unlist(pt1$lambda))
  # }
  # else {
  #   numdata[paste0(names.of.cols[i], ".pow")] <-
  #     -1 * ((numdata[i]) ^ as.numeric(unlist(pt1$lambda)))
  # }
  # generate and print test statistics

```

# finally, another way of arranging figures in a group using layout()

```{r fun with the layout function, fig.height=7, fig.width=10}
par(mar = c(3,4,1,2), mgp = c(1.7,0.3,0), tcl = 0.3, 
    font.lab = 2, cex.lab = 1.4, cex.axis = 1.2, lend = "square")
layout(matrix(c(1,2,1,2,3,3), byrow = TRUE, 
              nrow = 3, ncol = 2))
with(sv18, plot(Northing ~ Easting, 
                pch=c(15,19,17)[Type], col=c(2,4,6)[Type], 
                asp = 1, cex = 2))
with(sv18, text(Easting, Northing,
                labels = Group, pos = c(2,4),
                cex = 1, col = seq(2,10)[Group]))
grid()
legend("bottomleft", legend=levels(sv18$Type), 
       pch = c(15,19,17), col = c(2,4,6),
       cex = 1.4, pt.cex = 2,
       title = "numbers refer to Groups",
       title.col = 8, bty = "n")
with(sv18, boxplot(log10(Pb) ~ Type, 
                   ylab = expression(bold(paste(log[10],"(Pb, mg/kg)")))))
with(sv18, stripchart(pH ~ sv18$Group, method = "jitter", vertical = TRUE,
                      cex = 1.4, pch = 3, lwd = 2, col = seq(2,9), 
                      xlab = "Sampling group"))
box(lwd=2)
```

# Different map tiling packages

## coordinate conversion
```{r}
require(PBSmapping)
TEMP.utm=as.matrix(cbind(sv18$Easting, sv18$Northing))
colnames(TEMP.utm) <- c("X","Y")
attr(TEMP.utm, "projection") <- "UTM"
attr(TEMP.utm, "zone") <- 50
TEMP.utm
TEMP.LL <- convUL(TEMP.utm, km=FALSE, southern=TRUE)
TEMP.LL
sv18$Longitude <- TEMP.LL[,1]
sv18$Latitude <- TEMP.LL[,2]

rm(list = c("TEMP.utm", "TEMP.LL"))
```

## rosm package
```{r}
library(rosm)
library(prettymapr)

# > osm.types() # this is what rosm can do
# [1] "osm"              "opencycle"         "hotstyle"      "loviniahike"           
# [5] "loviniacycle"     "hikebike"          "hillshade"     "osmgrayscale"          
# [9] "stamenbw"         "stamenwatercolor"  "osmtransport"  "thunderforestlandscape"
# [13] "thunderforestoutdoors"  "cartodark"        "cartolight"  

par(mar = c(3,3,1,1), mgp = c(1.7,0.3,0), lend=2, tcl=0.25)

sv.rosm <- makebbox(-31.9298,115.8525,-31.9345,115.8476)
# just using plot functions in rosm
osm.plot(sv.rosm, type="hotstyle", zoomin=-1)
osm.points(sv18$Longitude,sv18$Latitude, 
           col="red3", lwd=2, pch=3, cex=1.5)

# embedding within prettymap()
prettymap({
  osm.plot(sv.rosm, type="stamenwatercolor", zoomin = -1)
  osm.points(sv18$Longitude,sv18$Latitude, 
             col=c(4,2,6)[sv18$Type], lwd=2, 
             pch=c(19,15,17)[sv18$Type], cex=1.5)
  osm.text(115.8507, -31.9324, 
           labels = "Smith's\nLake", 
           font = 3, col = "darkcyan")
  osm.text(115.850, -31.9308, 
           labels = "Charles Veryard\nReserve", 
           font = 3, col = "darkgreen")
},
  oma = c(3,3,1,1), drawarrow = T, arrow.scale = 1.5,
  scale.htin = 0.2, scale.label.cex = 1.5
  )
legend(par("usr")[1] + 0.075*(par("usr")[2]-par("usr")[1]),
       par("usr")[3] + 0.3*(par("usr")[4]-par("usr")[3]), 
       legend = levels(sv18$Type),
       pch = c(19,15,17), col = c(4,2,6),
       inset = 0.02, cex = 1.5, pt.cex = 1.5)
box()
require(OpenStreetMap)
xtix <- projectMercator(rep(sv.rosm["y","min"],
                            length(pretty(sv.rosm[1,])[-1])),pretty(sv.rosm[1,])[-1])
axis(1, cex.axis = 1.2, at = xtix[,1], 
     labels = round(pretty(sv.rosm[1,])[-1],3))
mtext("Longitude (\u00B0E)", 1, 1.7, font = 2, cex = 1.5)
ytix <- projectMercator(pretty(sv.rosm[2,]),
                        rep(sv.rosm["x","min"], length(pretty(sv.rosm[2,]))))
axis(2, cex.axis = 1.2, at = ytix[2:(length(pretty(sv.rosm[2,]))-1),2], 
     labels = round(pretty(sv.rosm[2,])[2:(length(pretty(sv.rosm[2,]))-1)],3))
mtext("Latitude (\u00B0S)", 2, 1.7, font = 2, cex = 1.5)
rm(list = c("xtix", "ytix"))
```

## ggmap package
```{r make ggmap object}
require(ggmap)
# make ggmap object
register_google(key = "AIzaSyDU7QiTWE4RGFFQNmhWy51n7e4RBeHKjc0")
sv.ggGmap2 <- get_googlemap(center=c(115.85,-31.9319), 
                            zoom = 17, 
                            maptype = "roadmap", 
                            color = "color")
```

```{r draw map, fig.height=8, fig.width=8}
# draw map
ggmap(sv.ggGmap2) + 
  geom_text(aes(x = 115.8505, y = -31.9326, label = "Smith's\nLake",
                fontface = "italic", family="sans"), 
            size = 8, vjust = 0, hjust = -0.5, color="lightblue") + 
  geom_point(data = sv18, aes(x =Longitude, y=Latitude, color=Type), shape=19, size=2) +
  scale_colour_manual(values = c("blue", "black", "red")) +
  geom_text(aes(x = 115.85095, y = -31.93315, label = "Outlet",
                fontface = "plain", family="sans"), 
            size = 8, vjust = 0, hjust = -0.5, color="white") + 
  geom_text(aes(x = 115.851, y = -31.9321, label = "Inlet",
                fontface = "plain", family="sans"), 
            size = 8, vjust = 0, hjust = -0.5, color="white") +
  labs( x = "Longitude (\u00B0E)", y = "Latitude (\u00B0S)") +
  theme(axis.text=element_text(size=11, color="black"),
        axis.title=element_text(size=14,face="bold"),
        legend.text=element_text(size=11),
        legend.title=element_text(size=12,face="bold"),
        legend.key=element_rect(fill="grey92"))
```

