---
title: "CyboRg 2021-09 (September)"
output: 
  word_document: 
    fig_width: 5
    fig_height: 5
---

# CyboRg session 2021-09-21

This is code to populate an irregular polygon with a selected number of randomized points. It represents part the first phase of development of an R package to generate different sampling designs given a site boundary and number or spacing of samples. In this case of randomized points, spacing is not really a relevant concept.

## read data and set up basic boundary parameters
```{r}
irreg <- read.csv("sv_boundary.csv")
str(irreg)
# use floor, ceiling to get integer values
extremes <- as.integer(c(min(floor(irreg$Easting)),
              min(floor(irreg$Northing)),
              max(ceiling(irreg$Easting)),
              max(ceiling(irreg$Northing))))
```

## use boundary data to set up and generate random points
```{r message=FALSE, warning=FALSE}
# we need the geoR package for the locations.inside() function
library(geoR)

# make x and y vectors with 1m spacing to randomly select from
xlist <- seq(extremes[1], extremes[3], 1)
ylist <- seq(extremes[2], extremes[4], 1)

# choose how many points within the irregular polygon...
np <- 60
#...and make an empty data frame to hold the results
inside <- data.frame(E = rep(NA, np), N = rep(NA, np))

# initiate a counting variable...
j = 1

# ...and use a while loop to check whether the random samples from
# the x and y vectors are inside the polygon, adding them to the data 
# frame if they are. Keep looping until the data frame is full.
while(j <= np) {
  data0 <- data.frame(E = sample(xlist, 1), N = sample(ylist, 1))
  check <- locations.inside(locations = data0,
                            borders = irreg[, c("Easting", "Northing")])
  if (class(check[1, 1]) == "numeric") {
    inside[j,] <- data0[1,]
    j <- j + 1
  }
}
str(inside)
```

## check how it looks with a plot

We use the objects irreg, extremes, inside from the previous 2 chunks.

```{r fig.height=5, fig.width=4}
par(mfrow = c(1,1), mar = c(3,3,1,1), mgp = c(1.6, 0.3, 0), 
    tcl = 0.25, font.lab = 2, lend=1, ljoin=2, cex.axis = 0.8)
with(irreg, plot(Easting, Northing, asp = 1, type = "l", cex.lab = 1.1))
rect(extremes[1], extremes[2], 
     extremes[3], extremes[4], 
     border = "tan", lty = 2)
points(inside[,1:2], pch = 3, col = "red3")
legend("bottomleft", bty = "o", inset = 0.06, cex = 0.9,
       box.col = "grey", bg = "white", y.intersp = 1.2,
       legend = c("Irregular polygon", "Max. extent", paste(np,"random points")),
       pch = c(NA,NA,3), col = c(1, "tan", "red3"), lty = c(1,2,NA))
```

### tidy up
```{r}
rm(list = c("xlist", "ylist", "np", "inside", "j", "data0", "check"))
```
